---
title: "ICCAT Bigeye Tuna assessment"
format: html
editor: visual
---

```{r}
#| label: install-jabba
#| include: false
remotes::install_github("jabbamodel/JABBA")
library(JABBA)
```

```{r}
#| label: file-setup
#| include: false
library(here)
file = here("Example/Example_iccat.R")
assessment = "BETiccat"
out.dir = here("Example","iccat", assessment)
if(!file.exists(out.dir) ) {
  dir.create(out.dir)
}
```

## ICCAT Bigeye Tuna assessment

```{r}
#| label: iccat-data
#| echo: false 
#------------------------------------------------------
# Simple example fit JABBA to Catch and CPUE with SEs
#-------------------------------------------------------
data(iccat)
# get BET data
bet = iccat$bet
```

```{r}
#| label: build-JAGS model
#| message: false
#| warning: false
#| cache: true
# Compile JABBA JAGS model and input object
jbinput1 = build_jabba(catch=bet$catch,cpue=bet$cpue,se=bet$se,
                      assessment=assessment,scenario = "Run1",
                      model.type = "Fox",
                      r.prior = c(0.2,0.5),
                      sigma.est = FALSE, # estimate additional observation error
                      fixed.obsE = 0.01, # mimum observation error
                      igamma = c(0.001,0.001), # uninformative inv-gamma for process error
                      psi.prior = c(1,0.2), # Initial depletion B/K
                      verbose=F)
# Check input
jbplot_indices(jbinput1)

# Fit JABBA (here mostly default value - careful)
bet1 = fit_jabba(jbinput1,quickmcmc = TRUE) # quick run
```

```{r}
#| label: indvidual-plots
head(bet1$kobe)

# Make individual plots
jbplot_catch(bet1)
jbplot_catcherror(bet1)
jbplot_ppdist(bet1)
jbplot_mcmc(bet1)
jbplot_residuals(bet1)
jbplot_cpuefits(bet1)
jbplot_runstest(bet1)
jbplot_logfits(bet1)
jbplot_procdev(bet1)
jbplot_PPC(bet1) # Posterior Predictive Checks - Not great should 0.2-0.8

# Status
jbplot_summary(bet1)

# combine plots
jbpar(mfrow=c(2,2))
jbplot_summary(bet1,add=T,type = c("BBmsy", "FFmsy"))
jbplot_spphase(bet1,add=T)
jbplot_kobe(bet1,add=T)

jbpar(mfrow=c(3,2),plot.cex = 0.8)
jbplot_ensemble(bet1)
```

```{r}
#| label: variance-model
#| message: false
#| warning: false
#| cache: true
# Try to improve runs test diagnostics by changing the variance settings
# Increase minimum obs error from 0.01 to 0.1 and remove SEs from CPUE model
jbinput2 = build_jabba(catch=bet$catch,cpue=bet$cpue,se=NULL,assessment=assessment,
                      scenario = "Run2",model.type = "Fox",sigma.est = TRUE,fixed.obsE = 0.1,igamma = c(0.001,0.001),
                      psi.prior = c(1,0.2), # Initial depletion B/K
                      verbose = F)


bet2 = fit_jabba(jbinput2,quickmcmc = T)
```

```{r}
#| label: variance-plots
# Check residual diags
jbplot_cpuefits(bet2)
jbplot_runstest(bet2)
jbplot_logfits(bet2)
jbplot_PPC(bet2)
# Improved
refinput = jbinput2 # Note as reference input 
# Compare
jbplot_summary(list(Run1=bet1,Run2=bet2))
jbplot_ensemble(list(Run1=bet1,Run2=bet2))

# Check parameters and convergence (p <0.05 is not fully converged)
bet2$pars 
```

```{r}
#| label: full-run model 
#| warning: false
#| message: false
#| cache: true
# Make a long MCMC run with 3 chains
bet.full = fit_jabba(jbinput2,nc=3)
```

```{r}
#| label: full-run values
# MCMC convergence
bet.full$pars 
jbplot_mcmc(bet.full)

# get quantaties
tibble::as_tibble(bet.full$estimates)
# FLR data.frame trajectories
tibble::as_tibble(bet.full$flqs)
# fits
tibble::as_tibble(bet.full$diags)


```

## Estimate shape m as function of Bmsy/K

```{r}
#| label: shape m model
#| message: false
#| warning: false
#| cache: true
# Compile JABBA JAGS model and input object
jbinput3 = build_jabba(catch=bet$catch,cpue=bet$cpue,se=NULL,assessment=assessment,
              scenario = "Est.Shape",model.type = "Pella_m", # Estimate shape
              BmsyK=0.4, # mean 40%B0
              shape.CV = 0.3, #CV
              sigma.est = TRUE,
              fixed.obsE = 0.1,
              igamma = c(0.001,0.001),
              psi.prior = c(1,0.1))

bet3 = fit_jabba(jbinput3,quickmcmc=F)
```

```{r}
#| label: shape-m checks
jbplot_ppdist(bet3) # check shape prior & posterior dist - not much information
# Compare
jbplot_summary(list(bet2,bet3))
jbplot_ensemble(list(bet2,bet3))
```

```{r}
#| label: Schaefer-model
#| message: false
#| warning: false
#| cache: true
# also run model as Schaefer
jbinput4 = build_jabba(catch=bet$catch,cpue=bet$cpue,se=NULL,assessment=assessment,
                       scenario = "Schaefer",model.type = "Schaefer", # Estimate shape
                       sigma.est = TRUE,
                       fixed.obsE = 0.1,
                       igamma = c(0.001,0.001),
                       psi.prior = c(1,0.1))

bet4 = fit_jabba(jbinput4,quickmcmc=T)
```

```{r}
#| label: Schaefer-plots
# Compare 
jbpar(mfrow=c(3,2),plot.cex=0.7)
jbplot_ensemble(list(bet1,bet2,bet3,bet4))
jbpar(mfrow=c(3,2),plot.cex=0.6)
jbplot_summary(list(bet1,bet2,bet3,bet4),add=T)
```

# Forecasting

## F-based Forecasting

```{r}
#| label: F-based forecasting
# F-based forecasting
# Relative Fmsy
# Single Forecast for Base-Case model - now works with imp.yr=1 
fw1 = fw_jabba(bet2,nyears=10,imp.yr=1,imp.values = seq(0.8,1.2,0.1),quant="F",type="msy",stochastic = T)
#jbpar(mfrow=c(3,2))
jbpar(mfrow=c(3,2),plot.cex = 0.7)
jbplot_ensemble(fw1)
# Zoom-in
jbplot_ensemble(fw1,xlim=c(2010,2027))
abline(v=2018) # Check
# Forecast with AR1 process error
fw1.ar1 = fw_jabba(bet2,nyears=10,imp.yr=1,quant="F",type="msy",AR1=TRUE,stochastic = T)
# now compare
jbpar(mfrow=c(3,2),plot.cex = 0.6)
for(i in 1:3){
  jbplot_ensemble(fw1,subplots = c(1,2,5)[i],add=T,xlim=c(2010,2028),legend=ifelse(i==1,T,F))
   jbplot_ensemble(fw1.ar1,subplots = c(1,2,5)[i],add=T,xlim=c(2010,2028),legend=ifelse(i==1,T,F))
   }
mtext(c("Default","AR1"),outer=T,at=c(0.27,0.77))
```

## IOTC style

```{r}
#| label: IOTC-style-forecast
# IOTC-Style: Relative current catch (default mean 3 yrs)
# 10 years, 2 intermediate years, deterministic
fw.io = fw_jabba(bet2,nyears=10,imp.yr=3,imp.values = seq(0.6,1.2,0.1),quant="Catch",type="ratio",nsq=3,stochastic = F)
jbplot_ensemble(fw.io)
jbpar(mfrow=c(2,2))
jbplot_ensemble(fw.io,add=T,subplots = 1,legend.loc = "topright")
jbplot_ensemble(fw.io,add=T,subplots = 2,legend=F)
jbplot_ensemble(fw.io,add=T,subplots = 5,legend=F)
jbplot_ensemble(fw.io,add=T,subplots = 6,legend=F)
```

## ICCAT style
```{r}
#| label: ICCAT-style-forecast
# ICCAT Style
Ccur = mean(tail(jbinput2$data$catch[,2],2))
TACs = c(75500,seq(60000,78000,2000))
fw.iccat= fw_jabba(bet2,nyears=10,imp.yr=3,initial = c(Ccur,76000),imp.values = TACs,quant="Catch",type="abs",nsq=3,stochastic = F,AR1=T)

jbpar(mfrow=c(2,2))
jbplot_ensemble(fw.iccat,legendcex = 0.4,xlim=c(2010,2027),subplots = c(1,2,5,6),add=T)
jbplot_ensemble(fw.iccat,legendcex = 0.4,xlim=c(2010,2027),subplots = c(1,2,5,6),plotCIs = F)

```

```{r}
#| label: verify-forecasts
# Check if correct
jbpar()
jbplot_ensemble(fw.iccat,legendcex = 0.4,xlim=c(2010,2027),subplots = c(6),plotCIs = F,add=T)
abline(v=c(2017,2018,2019,2020)) # 2020 = imp.yr
```

## Ensemble

```{r}
#| label: ensemble model forecast
# Do Ensemble modelling
jbplot_ensemble(list(bet2,bet3,bet4))

# Joint all runs
ens = jbplot_ensemble(list(bet2,bet3,bet4),kbout=T,joint=T)

# Do ensemble forecast

fw.ens= fw_jabba(list(bet2,bet3,bet4),nyears=10,imp.yr=2,initial = Ccur,imp.values = TACs,quant="Catch",type="abs",nsq=3,stochastic = F,AR1=T,thin=3)
jbpar(mfrow=c(3,2),plot.cex = 0.6)
for(i in 1:6) jbplot_ensemble(fw.ens,add=T,subplots = i,legend = ifelse(i==2,T,F))


```

# Conduct Retrospective Analysis and Hind-Cast Cross-Validation

## Hindcast cross validation
```{r}
#| label: hcxvl
# Do hindcast cross-validation
hc1 = hindcast_jabba(jbinput2,bet2,peels=1:5)

# Show Retrospective Pattern
mohns= jbplot_retro(hc1)

mohns
mohns[row.names(mohns)=="rho.mu",]

hindcasts = hc1
```

## Alternative Forecasts
```{r}
#| label: Alt-hindcast forecast
# Make alternative forecasts
hc2 = jbhcxval(hc1,AR1=T) # make forecasts with AR1

jbpar(mfrow=c(1,2))
for(i in 1:1){
  jbplot_hcxval(hc1,index=c(2)[i],add=T,minyr = 2000,legend.add = F)
  jbplot_hcxval(hc2,index=c(2)[i],add=T,minyr = 2000,legend.add = F)
}
mtext(c("Default","AR1"),outer=T,at=c(0.27,0.77))

jbmase(hc2)
```

# Catch-Only with biomass prior in 2010 as type B/Bmsy

```{r}
#| label: catch-only model
#| message: false
#| warning: false
#| cache: true
# Compile JABBA JAGS model and input object for Catch Only
# Add biomass prior based on B/Bmsy guestimate
jbinput5 = build_jabba(catch=bet$catch,model.type = "Fox",
                      assessment=assessment,scenario =  "CatchOnly" ,
                      b.prior=c(0.5,0.2,2010,"bbmsy"),
                      psi.prior = c(1,0.1))
# Fit JABBA
bet5 = fit_jabba(jbinput5,save.jabba=TRUE,output.dir=out.dir)
```

```{r}
#| label: catch-only-checks
# Check depletion prior vs posterior
jbplot_bprior(bet5)
# Compare
jbplot_summary(list(bet2,bet5))
```

# Original South Atlantic Swordfish example.Winker et al. (2018). JABBA: Just Another Biomass Assessment

```{r}
#| label: swos-setup
swos = iccat$swos

assessment = "SWOSiccat"
scenario = "Base"
out.dir = here("Example", "iccat", assessment)
if(!file.exists(out.dir) ) {
  dir.create(out.dir)
}
```

```{r}
#| label: swos-model
#| message: false
#| warning: false
#| cache: true
# Compile JABBA JAGS model and input object
jbswos = build_jabba(catch=swos$catch,cpue=normIndex(swos$cpue),se=swos$se,assessment=assessment,scenario = scenario,
                      model.type = "Pella",
                      BmsyK = 0.4,
                      r.prior=c(0.42,0.37),
                      K.prior = c(250000,1),
                      psi.prior = c(1,0.25),
                      fixed.obsE = 0.2,
                      add.catch.CV = FALSE,
                      proc.dev.all = FALSE, 
                      igamma=c(4,0.01), # default process error (moderately informative)
                      P_bound = c(0.02,1.1),verbose=FALSE)

```

```{r}
#| label: swos-indices
jbplot_indices(jbswos)
```
```{r}
#| label: swos-fits
#| message: false
#| warning: false
#| cache: true
# fit JABBA
fit.swos = fit_jabba(jbswos,save.jabba=TRUE,output.dir=out.dir)
```

```{r}
#| label: swos-plots
jbplot_cpuefits(fit.swos)
jbplot_logfits(fit.swos)
jbplot_residuals(fit.swos)
jbplot_runstest(fit.swos)
jbplot_ppdist(fit.swos)

jbpar(mfrow=c(1,2),plot.cex = 0.7)
jbplot_spphase(fit.swos,add=T)
jbplot_kobe(fit.swos,add=T)

```

```{r}
#| label: swos-iccat
# Project
# ICCAT Style
Ccur = 10056
TACs = seq(10000,18000,1000)
fw.swos= fw_jabba(fit.swos,nyears=10,imp.yr=4,initial = Ccur,imp.values = TACs,quant="Catch",type="abs",nsq=3,stochastic = T,AR1=F)
jbpar(mfrow=c(2,2),plot.cex = 0.7)
jbplot_ensemble(fw.swos,legendcex = 0.4,xlim=c(2015,2025),subplots = c(1:4),add=T)
```
